---
title: Using ggplot - Histograms
author: Ron Richardson
date: '2017-09-25'
slug: using-ggplot-histograms
categories:
  - DATA 650
  - R
tags:
  - charts
  - visualization
---

## Introduction
A histogram uses rectangular bars to represent data that is grouped into a series, sized accordingly based on the numerical value associated with it. The rectangular bars can be displayed either vertically or horizontally.

## Basic Example

First, we must gather our data.  Here we are using the Lahman database and obtaining all of the weights of baseball players throughout the years.

```{r message=FALSE, warning=FALSE}
#import libraries
library(Lahman)
library(dplyr)
library(ggplot2)

#store data into a variable named df
df<-Master %>%
  filter(!is.na(weight)) %>%
  select(weight)
```

To display the chart, we use the ggplot library. The 'data' parameter is the data source, and 'aes' refers to the aesthetics.

```{r}
ggplot()+
  geom_histogram(data=df,aes(x=weight), color="white", fill="blue", bins=25)+ 
  ggtitle("Baseball player weights")
```

