---
title: Dynamic Wordclouds with Wordcloud2
author: Ron Richardson
date: '2017-11-13'
slug: dynamic-wordclouds-with-wordcloud2
categories:
  - R
tags:
  - wordcloud
  - wordcloud2
---

## Introduction

[Wordclouds](/post/creating-a-wordcloud/) are great tools to quickly see the words used in a piece of text, but they can get boring to look at. In this post, we introduce a new R package called wordcloud2. This package will dynamically plot the wordcloud on screen, add hoverability to each word, and be able to form the wordcloud into different shapes.

## Getting our Data

Once again, we will use the text from Bram Stoker's "Dracula". For a more indepth explanation of how we get this data, please see the post ["Simple Sentiment Analysis"](/post/simple-sentiment-analysis/).

```{r message=FALSE, warning=FALSE}
library(gutenbergr)
library(tidytext)
library(dplyr)
library(wordcloud2)

dracula<-gutenberg_download(345)
dracula$gutenberg_id<-NULL

dracula<-dracula%>%
  unnest_tokens(word, text)

bing<-get_sentiments('bing')
dracula<-inner_join(dracula, bing)
```

We also need to count up the number of occurrences of each word in the text, using dplyr.

```{r message=FALSE, warning=FALSE}
words<-dracula%>%
  group_by(word)%>%
  summarize(count=n())%>%
  arrange(count)
```

## Using wordcloud2

Wordcloud2 has several options for how to shape the wordcloud. The available shapes are:
* circle
* cardioid
* diamond
* triangle
* triangle-forward
* pentagon
* star

For this example, we will make a triangle of the words, and also change the background color to be black.

```{r message=FALSE, warning=FALSE}
wordcloud2(words, shape='triangle', backgroundColor='black')
```


### The Code

The code for this post can be found on my [GitHub Gists page](https://gist.github.com/rer145/edb11422ad6ceffe220b9002fa4624d0).